---
title: "dimensional"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dimensional}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dimensional)
library(tidygraph)
library(igraph)
library(tidyverse)
```
# How to use the package:
## Data:
Dimensional takes data in a table graph format. This can be added a number of ways, for details see https://www.data-imaginist.com/2017/introducing-tidygraph/. For an example, we will here use an edgelist and a nodelist to create an undirected table graph.
```{r make tbl-graph}
decor.graph <- tbl_graph(nodes = nodes_decor, edges = edges_decor, directed = FALSE)

decor.graph
```
In this example data, the different dimensions of the graph are stored as edge attributes under the `name` variable. You should make sure that your data also stores the dimensions this way, and be certain to call this variable "name".

## `flatten()`
This function flattens a multilayer graph into a simple graph. This function is useful for comparing multilayer with simple network measures. 
`flatten()`simply collapses all edges between a pair of actors across all layers into a single edge. If there is a weight column, the weights of all collapsed edges will be summed. If weight = NULL, no weights are retained. 
```{r}
simple.flat <- flatten(decor.graph, weight = NULL)

simple.flat

weighted.flat <- flatten(decor.graph, weight = weight)

weighted.flat
```
## `multidesc()`: 
This function will give a particular set of measures for a basic description of a multilayer network, both layer by layer and for the flattened network. The measures are: number of nodes, density, clustering coefficient, degree centralization, average path length, diameter.
```{r multidesc}
multidesc(decor.graph)
```

## `centrality_neighborhood()`
This augments the `centrality_` classes in tidygraph with neighborhood centrality as defined by Berlingerio et al. (2011). `centrality_neighborhood()` can take a specified actor, or a vector of actors, as well as a specified layer of vector of layers. If these are left unspecified, it defaults to the entire graph. The function appends a neighborhood centrality measure as a node attribute named "neighbours"
```{r centrality_neighborhood}
a <- c("5", "6")  #specify actors by id
l <- c("folded_strip_roulette")

centrality_neighborhood(decor.graph)

```
## `connective_redundancy()``
Calculates the connective redundancy of a set of actors over a set of layers as defined by Dickison et al. (2016). Can take a specified actor, or a vector of actors, as well as a specified layer of vector of layers. If these are left unspecified, it defaults to the entire graph. Returns a tibble with actor name, degree, neighbors, and redundancy.
```{r connective redundancy}
conred <- connective_redundancy(decor.graph)

head(conred)
```

## `dimensional_correlation()``
```{r}
dimension.correlation <- dimension_correlation(decor.graph) 


#plotting a heatmap for the correlation matrix
dimension.correlation[lower.tri(dimension.correlation, diag = TRUE)] <- NA
cor_tri <- as.data.frame(dimension.correlation) %>% 
  mutate(Var1 = factor(row.names(.), levels=row.names(.))) %>% 
  gather(key = Var2, value = value, -Var1, na.rm = TRUE, factor_key = TRUE) 

ggplot(data = cor_tri, aes(Var2, Var1, fill = value)) + 
  geom_tile()+
 scale_fill_viridis_c()+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

